The Anatomy of a Temporal Summarization System

### 2015-01-24 introduction.tex
updated intro and motivation Not yet done with motivation

\section{Introduction}
\label{sec:introduction}
During crises, information is critical for first responders and those caught
in the event.  When the event is significant, as in the case of Hurricane
Sandy, the amount of information produced by traditional news outlets,
government agencies, relief organizations, and social media can vastly
overwhelm those trying to monitor the situation. Methods for identifying,
tracking, and summarizing events from text based input have been explored
extensively  (e.g.,
\cite{allan1998topic,Filatova&Hatzivassiloglou.04a,Wang&al.11}). However,
these experiments were not performed in the large and heterogeneous
environment of the modern web; there is still a need for robust and scalable 
methods for automatic summarization.

@@

In this paper, we detail our participation in the second year of the TREC
Temporal Summarization (TS) track. Broadly, TS systems are required to monitor
a time ordered corpus for information about a specific event and ``emit'' 
updates about that event. Under the current TREC formulation of this task,
successful TS systems must synthesize techniques from information retrieval,
 extractive multi-document summarization, and update summarization.
Specifically, we must select updates to emit based on three primary criterion:
relevance, salience, and redundancy.

As this was our first time participating in TS track, we focused mostly on 
second and third criterion, relying on the TREC supplied resources for 
document retrieval and filtering heuristics for relevance.
Our system predicts sentence salience in the context of a
large-scale event and integrates these predictions into
a clustering based multi-document summarization system. We fit a Gaussian 
Process (GP) regression model to predict sentence salience and use 
these predictions to bias the
formation of sentence clusters around more salient regions in the input space
using affinity propagation (AP) clustering.  AP uses the salience predictions
as well as pairwise similarities among input sentences to identify
\emph{exemplar} sentences, which we use as our summary output.  Our approach
differs from other methods of summarization that compute salience by pairwise
comparisons alone, ignoring features of importance that are intrinsic to the
sentences themselves.

The remainder of the paper is organized as follows. We begin with a review of related work in the information retrieval and multi-document summarization literature.  In section
~\ref{sec:background} we give an overview of the semantic similarity, 
Gaussian process, and affinity propagation algorithms that make up the bulk of
our TS system. 
Next we describe the data with which we build our various models.
Then in section~\ref{sec:approach} we give a high level sketch
of our system, and then explain each component in detail.

>>>>

\section{Introduction}
\label{sec:introduction}
During crises, information is critical for first responders and those caught
in the event.  When the event is significant, as in the case of Hurricane
Sandy, the amount of information produced by traditional news outlets,
government agencies, relief organizations, and social media can vastly
overwhelm those trying to monitor the situation. Methods for identifying,
tracking, and summarizing events from text based input have been explored
extensively
%KM - I think we should consider adding other event summarizers. 
(e.g.,
\cite{allan1998topic,Filatova&Hatzivassiloglou.04a,Wang&al.11}). However,
these experiments were not developed to handle streaming data from  the large and heterogeneous
environment of the modern web. Neither were they developed to address the
information needs that arise during  crisis
situations. 
%KM - this last sentence sounds somewhat negative.
%there is still a need for robust and scalable 
%methods for automatic summarization.

In this paper, we present an update summarization system to track disasters
across time. Our system predicts sentence salience in the context of a
large-scale event, such as a disaster,  and integrates these predictions into
a clustering based multi-document summarization system. We train a regression
model to predict sentence salience and use these predictions to bias the
formation of sentence clusters around more salient regions in the input space
using affinity propagation (AP) clustering.  AP uses the salience predictions
as well as pairwise similarities among input sentences to identify
\emph{exemplar} sentences, which we use as our summary output.  Our approach
differs from other methods of summarization that compute salience by pairwise
comparisons alone, ignoring features of importance that are intrinsic to the
sentences themselves.

In addition to the tight integration between clustering and salience
prediction, our approach also exploits knowledge about disaster to determine
salience. Thus, salience does not just represent importance within a set of
documents; it also represents both how typical this sentence is of the input event
type (i.e., disaster, hurricane, tornado) and whether it specifies information
about this particular disaster. We use a set of language models, one for each
disaster type, to measure typicality of the sentence for the current event type. We
use a feature that measures distance of mentioned location from the center of
the disaster to represent its likelihood of referring to the input disaster.

@@

The remainder of the paper is organized as follows. We begin with a review of related work in the information retrieval and multi-document summarization literature.  In section
~\ref{sec:background} we give an overview of the semantic similarity,
%KM - Chris can you update this so it reflects the contributions. We don't
%mention semantic similarity. We do mention salience and the disaster specific
%features. Perhaps we should mention the use of a new approach for semantic
%similarity? 
Gaussian process, and affinity propagation algorithms that make up the bulk of
our TS system. 
Next we describe the data with which we build our various models.
Then in section~\ref{sec:approach} we give a high level sketch
of our system, and then explain each component in detail.

### 2015-01-24 motivation.tex
updated intro and motivation Not yet done with motivation

>>>>

%KM If we are short on space, not sure how much this section is needed as it's
%somewhat repetitive of intro.

%KM - Chris, could you get some good references to work in the crisis
%informatics space.
Crisis informatics\cite{?} is dedicated to finding methods for sharing the
right information in a timely fashion with relief organizations during a major crisis. With the
increasing impact of climate change, the world is seeing an increase in
disasters such as hurricanes, tornadoes, flooding and typhoons, all of which
cause major damage and wreak havoc with food supplies, housing, and health
issues. Health epidemics, such as the ebola crisis in West Africa, also create
a need for timely information about where problems are greatest. Social and
political 
crises, such as the current situation in Syria, create similar needs for
humanitarian assistance.

At times of crisis, however, social media can overwhelm current information
systems with the quantity of information, much of which is irrelevant,
unnecessary detail, or out of date. Many approaches have focused on

%KM - PRobably should make a little more inclusive, there are graph based
%models that use methods such lexical similarity to build the graph,
%probabilistic methods (e.g., based on word probabilities, regression or
%topical signatures). There are also methods that use topic segmentation or
%modeling and select sentences for each topic.
